---
title: "Setup"
subtitle: "Experimentation with workflow"
author: "Einar Hjörleifsson"
date: "2018-11-09"
output: 
  html_document: 
    fig_height: 6
    fig_width: 9
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE)
```

## Cloning Muppet_HCR

The admb-source code and some templates for setups for plaice and haddock are available in [Husky's github repo](https://github.com/hoski/Muppet_HCR). The easiest way to get this on your computer is via the following command in a terminal window:

```{sh, eval = FALSE}
git clone https://github.com/hoski/Muppet_HCR.git
```

I suggest that you do the above from within your working directory where you intend to do you own simulation. Once cloned, you have the following content in the `Muppet_HCR`-directory:
```
Muppet_HCR/
├── HCR
│   ├── Had
│   │   ├── Files
│   │   │   ├── autsurveydata.dat
│   │   │   ├── autsurveypar.dat
│   │   │   ├── catchandstockdata.dat
│   │   │   ├── catchparameters.dat
│   │   │   ├── hadprognosis.dat.biorule
│   │   │   ├── hadprognosis.dat.bioruleAge
│   │   │   ├── hadprognosis.dat.opt
│   │   │   ├── likelihoodparameters.dat
│   │   │   ├── marsurveydata.dat
│   │   │   ├── marsurveypar.dat
│   │   │   ├── optim.dat.hockey
│   │   │   ├── outputparameters.dat
│   │   │   ├── ProgWts.dat
│   │   │   ├── ReadHad.r
│   │   │   └── stockparameters.dat
│   │   └── StdSettings
│   │       ├── icehad.dat.opt
│   │       └── icehad.dat.prog
│   ├── Model
│   │   ├── muppet.tpl
│   │   ├── readallmcmc.r
│   │   └── Read.r
│   └── Plaice
│       ├── Files
│       │   ├── catchandstockdata.dat
│       │   ├── catchparameters.dat
│       │   ├── likelihoodparameters.dat
│       │   ├── optim.dat.hockey
│       │   ├── outputparameters.dat
│       │   ├── plaiceprognosis.dat.biorule
│       │   ├── plaiceprognosis.dat.Frule
│       │   ├── plaiceprognosis.dat.opt
│       │   ├── ReadPlaice.r
│       │   ├── stockparameters.dat
│       │   ├── surveydata.dat
│       │   ├── surveypar.dat
│       │   └── totcatch.txt
│       ├── HockeyEstAcf
│       │   ├── iceplaice.dat.opt
│       │   └── iceplaice.dat.prog
│       └── README
└── README.md
```

The key directory is 'HCR/Model' that contains the ad-model builder code in the muppet.tpl file. Then there are starting templates for two stocks, 'HCR/Haddock' and 
'HCR/Plaice'. Now we do not want to do our simulaton work in the Muppet_HCR directory, treat it just as a copy of the latest version of Husky's admb-code for doing the simultions.

## Setting up an MSE-evaluation

What follows is a work-flow example on how to seed an mse-evaluation within ones working directory, here taking the plaice as an example. The example directory here is called `23_mse_2019`, where we have the following content:

```
23_mse_2019
├── 23_mse_2019.Rproj
├── Muppet_HCR
├──  ... see above
├── README.html
├── README.Rmd
├── run.sh
└── setup.sh
```

Note that above we have the cloned `Muppet_HCR`-directory within our mse-directory. It can though be anywhere on your computer, one would only need to change the path in the shell-scripts below accordingly.

We want to start off my copying the Model-directory and the Plaice setup-directories. Here it is done by running the following shell-commands:

```{sh, eval = FALSE}
cp -rp Muppet_HCR/HCR/Model .
cp -rp Muppet_HCR/HCR/Plaice/Files .
cp -rp Muppet_HCR/HCR/Plaice/HockeyEstAcf 0Base
```

The next step is to compile the admb-code from within the Model directory we just copied and then do some household cleaning of intermediate files:
```{sh, eval = FALSE}
cd Model
admb muppet
rm muppet.cpp
rm muppet.htp
rm muppet.obj
cd ..
```

Move the compiled admb-program into the 0Base-directory:
```{sh, eval = FALSE}
cp -p Model/muppet 0Base/.
```

If you are going to test a harvest-rate rule copy the biorule template into the base-directory:
```{sh, eval = FALSE}
cp -p Files/plaiceprognosis.dat.biorule 0Base/plaiceprognosis.dat
```

The directory structure within your project directory now looks something like this:
```
23_mse_2019
├── 0Base
│   ├── iceplaice.dat.opt
│   ├── iceplaice.dat.prog
│   ├── muppet
│   └── plaiceprognosis.dat
├── 23_mse_2019.Rproj
├── Files
│   ├── catchandstockdata.dat
│   ├── catchparameters.dat
│   ├── likelihoodparameters.dat
│   ├── optim.dat.hockey
│   ├── outputparameters.dat
│   ├── plaiceprognosis.dat.biorule
│   ├── plaiceprognosis.dat.Frule
│   ├── plaiceprognosis.dat.opt
│   ├── ReadPlaice.r
│   ├── stockparameters.dat
│   ├── surveydata.dat
│   ├── surveypar.dat
│   └── totcatch.txt
├── Model
│   ├── muppet
│   ├── muppet.tpl
│   ├── readallmcmc.r
│   └── Read.r
├── Muppet_HCR
├──      ... not printed, see above
├── README.html
├── README.Rmd
├── run.sh
└── setup.sh
```

The above steps you can expect to do only once in every mse-evaluation cycle.

## Running an assessment

The muppet-program is in a nutshell both a "historical" assessment program with an objective function and a short and long term prognosis program. If we are only interested in an assessment we can simply do: 
```{r}
basedir <- getwd()
setwd("0Base")
system("./muppet -ind iceplaice.dat.opt  -nox")
setwd(basedir)
```

We can read the deterministic results via:
```{r}
rbx <- 
  fishvice::read_separ("0Base", ".")
```

And plot the summary results via:
```{r}
library(tidyverse)
rby <- 
  rbx$rby %>% 
  mutate(oY = ifelse(is.na(oY), pY, oY)) %>% 
  select(year, rec = r, bio, ssb, fbar, hr, catch = oY) %>% 
  gather(variable, value, -year)

rby %>% 
  ggplot(aes(year, value)) +
  geom_line() +
  facet_wrap(~ variable, scale = "free_y") +
  expand_limits(y = 0) +
  geom_vline(xintercept = 2018.5) +
  labs(x = NULL, y = NULL)
```

The residuals are visualized via:
```{r, fig.height = 9}
rbx$rbya %>% 
  select(year, age, rC, rU1) %>% 
  gather(variable, value, -c(year, age)) %>% 
  ggplot(aes(year, value)) +
  geom_col() +
  facet_grid(age ~ variable)
```

## Running the simulations

### Seeding some mcmc-runs

Lets do some limited runs of 100 thousands, saving every 200 iteration and hence ending up with 500 iterations:
```{r}
setwd("0Base")
system("./muppet -ind iceplaice.dat.opt -nox -mcmc 100000 -mcsave 200 -mcscale")
# create the mcmc files
system("./muppet -mceval -ind iceplaice.dat.prog")
setwd(basedir)
```

### One harvest scenario

The biorule plaice template run above was setup such that it runs on a 22% harvest rate on Biomass4+ as can be seen when viewing the plaiceprognosis.dat file:
```{sh}
cat 0Base/plaiceprognosis.dat
```

We also observe that the Btrigger value is set to 5 kt, hopefully low enough such that it is never activated when setting the TAC (test by setting it to a lower value).

Lets load the rmuppet-package that provides some convenient functions that read in the results:
```{r}
library(rmuppet)
d <- 
  rmuppet:::gather_mcmc("0Base") %>% 
  mutate(rec = rec / 1e3) %>% 
  gather(variable, value, -c(iter, year)) %>% 
  filter(variable != "assCV") %>% 
  group_by(year, variable) %>% 
  rmuppet:::calc_percentiles() %>% 
  ungroup()
d %>% 
  ggplot(aes(year)) +
  facet_grid(variable ~ ., scale = "free_y") +
  geom_vline(xintercept = 2018, colour = "grey") +
  geom_ribbon(aes(ymin = q05, ymax = q95), fill = "red", alpha = 0.1) +
  geom_ribbon(aes(ymin = q10, ymax = q90), fill = "red", alpha = 0.2) +
  geom_ribbon(aes(ymin = q25, ymax = q95), fill = "red", alpha = 0.4) +
  geom_line(aes(y = q50), colour = "red") +
  geom_line(data = rby, aes(year, value)) +
  expand_limits(y = 0) +
  labs(x = NULL, y = NULL)
```

## Multiple harvest scenarios

To setup a muliple harvest scenario, like testing for different harvest rates we need to setup some loops of the process above. Here the bookeeping is done from within an R-loop.
```{r, eval = FALSE}
# overwrite package function - has a bug
source("R/set_prognosis.R")
source("R/one_run.R")

ctrlfile  <- paste(basedir, "Files/plaiceprognosis.dat.biorule",sep="/")
# put this into the function
file.exists(ctrlfile)

pars <-
  expand.grid(hr = c(0.001, 0.05, 0.10, seq(0.15, 0.30, by = 0.01)),
              Btrigger = 0.1,  # set to a very low value
              rec.rho  = 0.35,
              w.cv     = 0.08,
              ass.cv   = 0.2,
              ass.rho  = 0.7,
              mW.years = 10)
cn <- names(pars)

setwd("0Base")
res <- list()
for(i in 1:nrow(pars)) {
  
  set_prognosis(infile   = ctrlfile,
                outfile  = "plaiceprognosis.dat",
                HarvestRate       = pars$hr[i],
                AssessmentCorr  = pars$ass.rho[i],
                AssessmentCV   = pars$ass.cv[i],
                Btrigger = pars$Btrigger[i],
                RecrCorr  = pars$rec.rho[i],
                WeightCV     = pars$w.cv[i],
                Meanwtyears = pars$mW.years[i])
  
  one_run("iceplaice.dat.prog")

  res[[i]] <- gather_results()
  for(j in 1:length(cn)) res[[i]][,cn[j]] <- pars[i,cn[j]]

}

alldata <-
  bind_rows(res)
sumdata <-
  alldata %>%
  group_by(hr, Btrigger, rec.rho, mW.years, w.cv, ass.cv, ass.rho) %>%
  summarise_results(n = 5) %>% 
  ungroup()
HCRsettings <- readLines(ctrlfile)

save(list=c("HCRsettings","alldata","sumdata"), file = "HCRrun.rda")
save(list=c("HCRsettings","sumdata"), file = "HCRrunsum.rda")

setwd(basedir)
```

## Some results

```{r, fig.heigth = 9}
load("0Base/HCRrun.rda")
d <-
  alldata %>%
  as_tibble() %>%
  mutate(hr2 = catch/refbio1) %>% 
  select(hr, Btrigger, year,
         catch, ssb, rec, fbar = refF, hr2, bio = refbio1) %>%
  mutate(rec = rec / 1000) %>%
  gather(variable, value, -c(hr, Btrigger, year)) %>%
  group_by(hr, Btrigger, year, variable) %>%
  rmuppet:::calc_percentiles() %>% 
  ungroup()

d %>%
  filter(hr %in% c(0.1, 0.15, 0.20, 0.25, 0.30)) %>% 
  ggplot(aes(year)) +
  theme_bw() +
  geom_vline(xintercept = 2018, colour = "grey") +
  geom_ribbon(aes(ymin = q05, ymax = q95),
              fill = "red", alpha = 0.2) +
  geom_ribbon(aes(ymin = q10, ymax = q90),
              fill = "red", alpha = 0.2) +
  geom_ribbon(aes(ymin = q25, ymax = q75),
              fill = "red", alpha = 0.2) +
  geom_line(aes(y = q50), colour = "red") +
  facet_grid(variable ~ hr, scale = "free_y") +
  expand_limits(y = 0)
```


```{r}
sumdata %>% 
  select(hr, mean = catchmean, median = catchmed, p10 = catch10) %>% 
  gather(variable, value, -hr) %>% 
  ggplot(aes(hr, value, colour = variable)) +
  geom_line() +
  geom_point() +
  labs(title = "Catch")

Blim <- 
  fishvice::read_separ("0Base", ".")$rby %>% 
  filter(year < 2015) %>% 
  summarise(Blim = min(ssb)) %>% 
  pull(Blim)

sumdata %>% 
  select(hr, mean = ssbmean, p05 = ssb05) %>% 
  gather(variable, value, -hr) %>% 
  ggplot(aes(hr, value, colour = variable)) +
  geom_hline(yintercept = Blim) +
  geom_line() +
  geom_point() +
  labs(title = "SSB")
```



```{r, echo = FALSE, eval = FALSE}
# some trial
library(tidyverse)
#devtools::install_github("einarhjorleifsson/Muppet_HCR", ref = "trial")
library(MuppetHCR)
library(rmuppet)

name <- "0Base"
# Setup directories with files
system(paste0("cp -rp ", system.file("Plaice/Files", package = "MuppetHCR"), " ."))
system(paste0("cp -rp ", system.file("Plaice/HockeyEstAcf", package = "MuppetHCR"), " ", name))
system(paste0("cp -rp ", system.file("Model", package = "MuppetHCR"), " ."))
devtools::session_info()
system(paste0("cp -p Model/muppet.tpl ", name, "/."))
system(paste0("cp -p Files/plaiceprognosis.dat.biorule ", name, "/plaiceprognosis.dat"))
```

